---
title: "Advanced Load Forecasting"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r chunks, include=FALSE}
#set up chunk options
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), 
                      tidy=FALSE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r libraries, include=FALSE}
#retrieve libraries
library(here)
library(ggplot2)
library(forecast)
library(cowplot)
library(Kendall)
library(tseries)
library(kableExtra)
library(dplyr)
library(smooth)
```

# Preparation

## Upload

```{r include=FALSE}
#upload test and training data
load_daily <- read.csv("./Data/Processed/dailyload.csv")
load_hourly <- read.csv("./Data/Processed/hourlyload.csv")
load_daily_train <- read.csv("./Data/Processed/dailyload_train.csv")
load_daily_test <- read.csv("./Data/Processed/dailyload_test.csv")

#create msts
load_daily_msts <- msts(load_daily$daily_average, 
                            seasonal.periods =c(7,365.25),
                            start=c(2005,1,1)) 

load_hourly_msts <- msts(load_hourly$load, 
                            seasonal.periods =c(24,168,8766),
                            start=c(2005,1,1,1)) %>% #is this the right hour?
                         na.interp()

load_daily_train_msts <- msts(load_daily_train,
                              seasonal.periods =c(7,365.25),
                              start=c(2005,1,1))

load_daily_test_msts <- msts(load_daily_test,
                             seasonal.periods =c(7,365.25),
                             start=c(2005,1,1))
```

## ACF & PACF

```{r}
Acf(load_daily_msts, lag = 20)
Pacf(load_daily_msts, lag = 20)
```

# Forecasts

## ARIMA

Since this model uses complex seasonalities, the ARIMA with Fourier terms is the only ARIMA model used for this dataset.

### Training Data Set

```{r echo=FALSE}
#autofit the arima model
daily_arima_train <- auto.arima(load_daily_train_msts,
                                   xreg = fourier(load_daily_train_msts, 
                                                  K = c(2, 5)), 
                                   seasonal = FALSE)

#create the arima forecast with the autofit
forecast_daily_arima_train <- forecast(object = daily_arima_train,
                                   xreg=fourier(load_daily_train_msts,
                                                K=c(2,5),
                                                h=50),
                                   h=59)


#write into csv
write.csv(forecast_daily_arima_train, 
         file = "~/Duke_R/TSA_Forecasting/Forecasts/Training/arima.csv",
         row.names = FALSE)
```

### Full Data set

```{r echo=FALSE}
#autofit the arima model
daily_arima_autofit <- auto.arima(load_daily_msts,
                                   xreg = fourier(load_daily_msts, 
                                                  K = c(2, 5)), 
                                   seasonal = FALSE)
print(daily_arima_autofit)

#create the arima forecast with the autofit
forecast_daily_arima <- forecast(object = daily_arima_autofit,
                                   xreg=fourier(load_daily_msts,
                                                K=c(2,5),
                                                h=59),
                                   h=59)

#plot the forecast
plot(forecast_daily_arima)

#write into csv
write.csv(forecast_daily_arima, 
         file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_daily_arima.csv",
         row.names = FALSE)
```

Series: load_daily_msts Regression with ARIMA(2,1,1) errors

Coefficients: ar1 ar2 ma1 S1-7 C1-7 S2-7 C2-7 S1-365 C1-365 S2-365 0.8923 -0.2585 -0.9732 73.7011 22.0578 21.6431 -26.4128 37.7803 164.1901 648.0535 s.e. 0.0273 0.0251 0.0231 25.4128 25.4048 10.8776 10.8791 91.5511 76.4274 57.8418 C2-365 S3-365 C3-365 S4-365 C4-365 S5-365 C5-365 774.5907 -58.3548 -24.8293 -69.0629 -53.3360 7.1060 -51.5052 s.e. 52.0811 49.1096 46.1556 45.5895 43.8778 43.9148 42.7590

sigma\^2 = 254493: log likelihood = -16729.35 AIC=33494.7 AICc=33495.01 BIC=33597.15

### Hourly Code

```{r, eval=FALSE}
#autofit the arima model
hourly_arima_autofit <- auto.arima(load_hourly_msts,
                                   xreg = fourier(load_hourly_msts, 
                                                  K = c(2, 2, 5)), 
                                   seasonal = FALSE)
print(hourly_arima_autofit)

#create the arima forecast with the autofit
forecast_hourly_arima <- forecast(object = load_hourly_msts,
                                   xreg=fourier(load_hourly_msts,
                                                K=c(2,2,5),
                                                h=59*24),
                                   h=59*24)

#plot the forecast
plot(forecast_hourly_arima)

#write into csv
write.csv(forecast_hourly_arima, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_hourly_arima.csv",
          row.names = FALSE)
```

## ETS + STL

### Training Data Set

```{r}
#create forecast
forecast_daily_ets_train <- stlf(load_daily_train_msts,
                           h=50)

#write into csv
write.csv(forecast_daily_ets_train, 
         file = "~/Duke_R/TSA_Forecasting/Forecasts/Training/ets.csv",
         row.names = FALSE)
```

### Full Data Set

```{r}
forecast_daily_ets <- stlf(load_daily_msts,
                           h=59)

#plot the forecast
plot(forecast_daily_ets)

#write into csv
write.csv(forecast_daily_ets, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_daily_ets.csv",
          row.names = FALSE)
```

### Hourly

```{r}
forecast_hourly_ets <- stlf(load_hourly_msts,
                           h=59*24)

#plot the forecast
plot(forecast_hourly_ets)

#write into csv
write.csv(forecast_hourly_ets, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_hourly_ets.csv",
          row.names = FALSE)
```

##TBATS 
### Training Data Set

```{r}
#create fit
daily_tbats_fit_train <- tbats(load_daily_train_msts)

#forecasting data
forecast_daily_tbats <- forecast(daily_tbats_fit_train, 
                           h = 50)
#write into csv
write.csv(forecast_daily_tbats, 
         file = "~/Duke_R/TSA_Forecasting/Forecasts/Training/tbats.csv",
         row.names = FALSE)
```

### Full Data Set

```{r}
daily_tbats_fit <- tbats(load_daily_msts)

#forecasting test data
forecast_daily_tbats <- forecast(daily_tbats_fit, 
                           h = 59)

#visualization
plot(forecast_daily_tbats)

#write into csv
write.csv(forecast_daily_tbats, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_daily_tbats.csv",
          row.names = FALSE)
```

### Hourly Data Set

```{r, eval=FALSE}
#fit TBATS
hourly_tbats_fit <- tbats(load_hourly_msts)

#forecasting test data
forecast_hourly_tbats <- forecast(hourly_tbats_fit, 
                           h = 59*24)

#visualization
plot(forecast_hourly_tbats)

#write into csv
write.csv(forecast_hourly_tbats, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_hourly_tbats.csv",
          row.names = FALSE)
```

## Neural Network

Let NN model pick the p and P values automatically by omitting them. Find best K values as below:

```{r, eval=FALSE}
# Set max K values to test
max_K1 <- 3  # For weekly seasonality
max_K2 <- 5  # For yearly seasonality

results <- expand.grid(K1 = 1:max_K1, K2 = 1:max_K2)
results$AICc <- NA

# Loop over all K1-K2 combinations
for (i in 1:nrow(results)) {
  Kvals <- c(results$K1[i], results$K2[i])
  xreg <- fourier(load_daily_msts, K = Kvals)
  
  fit <- auto.arima(load_daily_msts, xreg = xreg, seasonal = FALSE)
  results$AICc[i] <- fit$aicc
}

# Find best K1-K2 pair
best_row <- results[which.min(results$AICc), ]
cat("Best K values: K1 =", best_row$K1, ", K2 =", best_row$K2, "\n")
```

### Training Data Set

```{r}
fit_nn_daily_train <- nnetar(as.numeric(load_daily_train_msts))

forecast_daily_nn_train <- forecast(fit_nn_daily_train, 
                   h=50)

#write into csv
write.csv(forecast_daily_nn_train, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Training/nn.csv",
          row.names = FALSE)
```

### Full Data Set

```{r}
fit_nn_daily <- nnetar(as.numeric(load_daily_msts))

forecast_daily_nn <- forecast(fit_nn_daily, 
                   h=59)

#write into csv
write.csv(forecast_daily_nn, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_daily_nn.csv",
          row.names = FALSE)
```

### Hourly Data Set

```{r, eval=FALSE}
fit_nn_hourly <- nnetar(load_hourly_msts,
                      xreg = fourier(load_hourly_msts, K = c(2, 5)))

forecast_daily_nn <- forecast(fit_nn_daily, 
                   h=59*24,
                   xreg= fourier(load_daily_msts, 
                                 K = c(2, 5)))

#write into csv
write.csv(forecast_daily_nn, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_daily_nn.csv",
          row.names = FALSE)
```

## State Space - ES

### Training Data Set

```{r}
fit_SSES_daily_train <- es(load_daily_train_msts,
                           model="ZZZ",
                           h=50,
                           holdout=FALSE)

forecast_daily_SSES_train <-forecast(fit_SSES_daily_train,
                                     h = 50, 
                                     interval="prediction")

#write into csv
write.csv(forecast_daily_SSES_train, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Training/sses.csv",
          row.names = FALSE)
```

### Full Data Set

```{r}
fit_SSES_daily <- es(load_daily_msts,
           model="ZZZ",
           h=59,
           holdout=FALSE)

forecast_daily_SSES <-forecast(fit_SSES_daily,
                    h = 59, 
                    interval="prediction")

#write into csv
write.csv(forecast_daily_SSES, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_daily_SSES.csv",
          row.names = FALSE)
```

### Hourly Data Set

```{r, eval=FALSE}
fit_SSES_hourly <- es(load_hourly_msts,
                      model="ZZZ",
                      h=59,
                      holdout=FALSE)

forecast_hourly_SSES <-forecast(fit_SSES_hourly,
                                h = 59*24, 
                                interval="prediction")

#write into csv
write.csv(forecast_hourly_SSES, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/Raw/forecast_hourly_SSES.csv",
          row.names = FALSE)
```

## State Space - BSM

This model cannot handle multiple seasonalities and was excluded from our analysis.
