---
title: "Models Rosie"
author: "Rosie Wu"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
#set up chunk options
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), 
                      tidy=FALSE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(here)
library(ggplot2)
library(forecast)
library(tidyverse)
```

```{r}
#create full, training, and test datasets
dailyload <- read.csv("./Data/Processed/dailyload.csv")

dailyload_train <- filter(dailyload, date <= "2009-12-31")

dailyload_test <- filter(dailyload, date > "2009-12-31" & date < "2010-02-28")

#create timeseries
daily_msts <- msts(dailyload[,2],
                   start = c(2005,01,01), 
                   seasonal.periods =c(7,365.25))

daily_train_msts <- msts(dailyload_train[,2],
                         start = c(2005,01,01), 
                         seasonal.periods =c(7,365.25))

daily_test_msts <- msts(dailyload_test[,2],
                        start = c(2009,12,31), 
                        seasonal.periods =c(7,365.25))
```


## Seasonal Data
### ARMA models, 


### TBATs models
```{r}
model_tbats <- tbats(daily_train_msts)
fc_tbats <- forecast(model_tbats)
accuracy(fc_tbats, daily_test_msts)

```



### ETS Models


### Geometric Mean

Create a forecast using the training data and test it on the testing data.

```{r}
#geometric mean on original data
forecast_daily_mean_test <- meanf(daily_train_msts, h=59)
plot(forecast_daily_mean_test)
```

Check the scores of the training model.

```{r}
#residuals and scores for seasonal naive model
residuals_daily_mean <- checkresiduals(forecast_daily_mean_test)
score_daily_mean <- accuracy(forecast_daily_mean_test, dailyload_test[,2])
score_daily_mean
```

Train a new model on the full dataset and then write it into a csv.

```{r}
#train seasonal naive on original data
forecast_daily_mean <- meanf(daily_msts, h=59) 

#create dataframe of forecast
forecast_daily_mean_df <- data.frame(forecast_daily_mean) %>%
    mutate("date" = seq(as.Date("2011/01/01"), by = "day", length.out = 59)) %>%
    mutate("load" = round(Point.Forecast))%>%
    select(date, load) %>%
    `rownames<-`(NULL) 

#write into csv
write.csv(forecast_daily_mean_df, 
          file = "~/Duke_R/TSA_Forecasting/Forecasts/forecast_daily_mean.csv",
       row.names = FALSE)
```

### Seasonal Naive Model

Create a forecast using the training data and test it on the testing data.

```{r}
#Seasonal naive on original data
forecast_daily_snaive_test <- snaive(daily_train_msts, h=59)
plot(forecast_daily_snaive_test)
```

Check the scores of the training model.

```{r}
#residuals and scores for seasonal naive model
residuals_daily_snaive <- checkresiduals(forecast_daily_snaive_test)
score_daily_snaive <- accuracy(forecast_daily_snaive_test, dailyload_test[,2])
score_daily_snaive
```

Train a new model on the full dataset and then write it into a csv.

```{r}
#train seasonal naive on original data
forecast_daily_snaive <- snaive(daily_msts, h=59) 

#create dataframe of forecast
forecast_daily_snaive_df <- data.frame(forecast_daily_snaive) %>%
    mutate("date" = seq(as.Date("2011/01/01"), by = "day", length.out = 59)) %>%
    mutate("load" = round(Point.Forecast))%>%
    select(date, load) %>%
    `rownames<-`(NULL) 

#write into csv
write.csv(forecast_daily_snaive_df, 
          file = "./Data/Forecasts/forecast_daily_new.csv",
          row.names = FALSE)
```
